SUS 2.0: Updating the System Usability Scale to conform with insights from questionnaire design research

### 2013-09-14 paper_draft1.Rnw
replaced t-tests with WM tests, wrote discussion and conclusion

Table \ref{tab:acqui} provides means, standard deviations, and significance levels from Welch two-sample tests (which does not assume equal sample variances like the t-test). Given the much larger sample size of responses on system A than B, we observe more significant differences for comparisons for system A than B. Most notably, highly significant acquiescence bias at $p<0.05$ was found for 6 out of 10 items and the overall SUS score for system A. This result clearly shows that the SUS is significantly biased.

\begin{table}[h]
\small
\centering
\caption{Strong evidence for acquiescence bias in the SUS for individual items and the overall score}
\label{tab:acqui}
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
\# & M & SD & M & SD & p value \\
\midrule
1 & 8.52 & 2.12 & 6.69 & 2.98 & \textless0.001\\
2 & 9.09 & 1.85 & 8.15 & 2.47 & \textless0.001\\
3 & 8.58 & 1.79 & 8.90 & 1.83 & 0.008\\
4 & 8.19 & 2.36 & 8.66 & 1.68 & \textless0.001\\
5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.67\\
6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.49\\
7 & 8.61 & 1.68 & 8.90 & 1.79 & 0.012\\
8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.13\\
9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.056\\
\midrule
 & 84.46 & 12.91 & 82.33 & 11.39 & 0.010\\
\bottomrule
\end{tabular}
\end{table}
<<eval=FALSE>>=
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
@

@@

Our goal was to update the SUS items for future use given that the SUS is clearly the most established usability scale. These updated items with corresponding answer scales are presented in Table \ref{tab:sus2items}. For our evaluation of the SUS 2.0, we were unable to use the full 10 items and opted for using a reduced number of items which we refer to as the short SUS 2.0. The short SUS 2.0 consists of 5 items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (leanability, confidence, ease of use, consistency).

[We gotta talk about how this impacts the SUS score calculations, comparison to past scores, etc.]

@@

The two online courses that are compared were offered on two distinct online platforms that shared the same core features but differed considerably in design. Based on Nielsen's heuristic evaluation of the two systems, it was determined that one system had more usability problems than the other. The following criteria were used [go through these and say which ones were different/better etc.]

Visibility of system status:
%The system should always keep users informed about what is going on, through appropriate feedback within reasonable time.

Match between system and the real world:
%The system should speak the user's language, with words, phrases and concepts familiar to the user, rather than system-oriented terms. Follow real-world conventions, making information appear in a natural and logical order.

User control and freedom:
%Users often choose system functions by mistake and will need a clearly marked "emergency exit" to leave the unwanted state without having to go through an extended dialogue. Support undo and redo.

Consistency and standards:
%Users should not have to wonder whether different words, situations, or actions mean the same thing. Follow platform conventions.

Error prevention:
%Even better than good error messages is a careful design which prevents a problem from occurring in the first place. Either eliminate error-prone conditions or check for them and present users with a confirmation option before they commit to the action.

Recognition rather than recall:
%Minimize the user's memory load by making objects, actions, and options visible. The user should not have to remember information from one part of the dialogue to another. Instructions for use of the system should be visible or easily retrievable whenever appropriate.

Flexibility and efficiency of use:
%Accelerators—unseen by the novice user—may often speed up the interaction for the expert user such that the system can cater to both inexperienced and experienced users. Allow users to tailor frequent actions.

Aesthetic and minimalist design:
%Dialogues should not contain information which is irrelevant or rarely needed. Every extra unit of information in a dialogue competes with the relevant units of information and diminishes their relative visibility.

Help users recognize, diagnose, and recover from errors:
%Error messages should be expressed in plain language (no codes), precisely indicate the problem, and constructively suggest a solution.

Help and documentation:
%Even though it is better if the system can be used without documentation, it may be necessary to provide help and documentation. Any such information should be easy to search, focused on the user's task, list concrete steps to be carried out, and not be too large.

Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for two online course interfaces where one interface has an objectively higher usability level than the other. As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method. While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity. A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0).
<<eval=FALSE>>=
wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]))
wilcox.test(convSUS(m.sus2[,2:5]), convSUS(a.sus2[,2:5]))
@

@@

@@

\section{Acknowledgments}
We are grateful to the instructors of the two Stanford courses who allowed us conduct survey experiments in their post-course surveys.

@@

\begin{table*}
\small
\centering
\caption{Items from the proposed SUS 2.0 with answer scales}
\label{tab:sus2items}
\begin{tabular}{| l | p{7cm} | p{9cm} |}
\hline
\textbf{\#}  & \textbf{Questions} & \textbf{Answer Scales} \\
\hline
1$^\star$  & How much more is there to learn about the system? & Nothing at all, A little, A moderate amount, A lot, A great deal \\ \hline
2  & How likely are you to need the support of a technical person to be able to use the system? & \{Extremely, Very, Somewhat\} unlikely, Neither likely nor unlikely, \{Somewhat, Very, Extremely\} likely \\ \hline
3$^\star$  & How confident are you using the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} confident \\ \hline
4  & How cumbersome is it to use the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} cumbersome \\ \hline
5$^\star$  & How easy or difficult is it to learn how to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
6  & How complex is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} complex \\ \hline
7$^\star$  & How easy or difficult is it to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
8  & How integrated are the system’s various functions? & \{Not at all, Slightly, Moderately, Very, Extremely\} integrated \\ \hline
9$^\star$  & How consistent is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} consistent \\ \hline
10 & How much do you like or dislike the system? & \{Extremely, Moderately, Slightly\} dislike, Neither like nor dislike, \{Slightly, Moderately, Extremely\} like\\
\hline
\end{tabular}
\smallskip
\scriptsize{$^\ast$Items included in the Short SUS 2.0}
\end{table*}

>>>>

Table \ref{tab:acqui} provides means, standard deviations, and p-values from non-parametric Mann-Whitney tests of the hypothesis that there is no location shift (this non-parametric alternative of the t-test is used as scores are not normally distributed). We find highly significant differences at 99\% significance in all but two items and the overall SUS score. This is very strong evidence for the claim that the original SUS induces acquiescence bias.

\begin{table}[h]
\small
\centering
\caption{Strong evidence for acquiescence bias in the SUS for individual items and the overall score}
\label{tab:acqui}
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
\# & M & SD & M & SD & p value \\
\midrule
1 & 8.52 & 2.12 & 6.69 & 2.98 & \textless0.001\\
2 & 9.09 & 1.85 & 8.15 & 2.47 & \textless0.001\\
3 & 8.58 & 1.79 & 8.90 & 1.83 & \textless0.001\\
4 & 8.19 & 2.36 & 8.66 & 1.68 & 0.17\\
5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.86\\
6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.006\\
7 & 8.61 & 1.68 & 8.90 & 1.79 & \textless0.001\\
8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.01\\
9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.002\\
\midrule
 & 84.46 & 12.91 & 82.33 & 11.39 & \textless0.001\\
\bottomrule
\end{tabular}
\end{table}
<<eval=FALSE>>=
# testItems=function(i, oa, ra){
#     print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
# }
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
@

@@

Our goal was to update the SUS items for future use given that the SUS is clearly the most established usability scale. These updated items with corresponding answer scales are presented in Table \ref{tab:sus2items}. For our evaluation of the SUS 2.0, we were unable to use the full 10 items and opted for using a reduced number of items which we refer to as the short SUS 2.0. The short SUS 2.0 consists of five items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (confidence, ease of use, consistency, leanability).

The score calculation for the SUS 2.0 is different to that of the original SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale. The SUS 2.0 can be scored as a proportion of the maximum score that is achievable. First, assign values from 0--4 or 0--6 depending on the number of scale points such that 0 reflects the worse response and 4 or 6 the best. Second, sum up the values for all ten responses; the sum should be an integer between 0--48. Third, divide by 48 to get the SUS 2.0 score. For the short SUS 2.0, follow the same steps except that the sum of response values will lie between 0--24 and you divide by 24, not 48.

@@

The two systems offered the same basic features, browsing and playing video lectures, but differed considerably in their design. We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability. While both systems showed generally high usability, one system was deemed superior in these four categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation. This informal usability comparison was the basis for labeling one system as having `high usability' and the other `low usability'.

Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for the high-usability and the low-usability system. As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method. While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity. A Mann-Whitney test of the difference between the usability estimates for the two systems for each scale supports this result ($W$=18585, $p$=0.07 for the original SUS; $W$=74598, $p\textless$0.001 for the short SUS 2.0).
<<eval=FALSE>>=
wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]))
wilcox.test(convSUS(m.sus2[,2:5]), convSUS(a.sus2[,2:5]))
@

@@

The paper's unique contribution is to analyze the SUS from a questionnaire design perspective which has not been done before. We provide strong evidence that the original SUS induces acquiescence bias and address this and other potential issues by proposing the SUS 2.0: an updated version of the original SUS with statements rephrased as questions with relevant answer scales. The SUS 2.0 is less likely to induce response biases, because it conforms to insights from questionnaire design research. Moreover, we find that the short SUS 2.0, with only five items, is a more sensitive measure of usability than the original, ten-item SUS.

Future work could investigate the strength of association between the original SUS and the SUS 2.0, which was not possible in these studies as each respondent saw just one usability scale. Moreover, our findings are limited in that we could only administer the short SUS 2.0, which constraint us from performing an investigation of the SUS 2.0's factor structure.

@@

This paper presents compelling evidence that highlights big questionnaire design deficits in the SUS. The authors propose a redesign of the scale (the SUS 2.0) and provide evidence that a sub-scale of the SUS 2.0 has high reliability, concurrent validity, an single-factor structure, and better sensitivity than the orignal SUS. While more work on evaluating the SUS 2.0 is needed, the authors strongly recommend the use of the updated scale for measuring usability.

\section{Acknowledgments}
We are grateful to the instructors of the two Stanford courses for allowing us to conduct survey experiments in their post-course surveys.

@@

\begin{table*}
\small
\centering
\caption{SUS 2.0 questions with answer scales}
\label{tab:sus2items}
\begin{tabular}{| l | p{7cm} | p{9cm} |}
\hline
\textbf{\#}  & \textbf{Questions} & \textbf{Answer Scales} \\
\hline
1$^\star$  & How much more is there to learn about the system? & Nothing at all, A little, A moderate amount, A lot, A great deal \\ \hline
2  & How likely are you to need the support of a technical person to be able to use the system? & \{Extremely, Very, Somewhat\} unlikely, Neither likely nor unlikely, \{Somewhat, Very, Extremely\} likely \\ \hline
3$^\star$  & How confident are you using the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} confident \\ \hline
4  & How cumbersome is it to use the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} cumbersome \\ \hline
5$^\star$  & How easy or difficult is it to learn how to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
6  & How complex is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} complex \\ \hline
7$^\star$  & How easy or difficult is it to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
8  & How integrated are the system’s various functions? & \{Not at all, Slightly, Moderately, Very, Extremely\} integrated \\ \hline
9$^\star$  & How consistent is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} consistent \\ \hline
10 & How much do you like or dislike the system? & \{Extremely, Moderately, Slightly\} dislike, Neither like nor dislike, \{Slightly, Moderately, Extremely\} like\\
\hline
\end{tabular}
\smallskip
\scriptsize{$^\ast$Items included in the Short SUS 2.0}
\end{table*}
